---
output:
  github_document:
    pandoc_args: --webtex
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# EstimationTools

<!-- badges: start -->
<!-- badges: end -->

The goal of `EstimationTools` is to provide routines for parameter estimation of probability density/mass functions.

## Installation

<!-- You can install the released version of EstimationTools from [CRAN](https://CRAN.R-project.org) with: -->

You can install the released version of `EstimationTools` from [github](https://github.com/Jaimemosg/EstimationTools) with:

``` r
if (!require('devtools')) install.packages('devtools')
devtools::install_github('Jaimemosg/EstimationTools', force=TRUE)
```

## Example

This is a basic example which shows you how to solve a common problem:

$$
\begin{aligned} 
X &\sim N(\mu, \:\sigma^2) \\
\mu &= 160 \\
\sigma &= 6
\end{aligned}
$$

The solution for a data set generated with size $n=10000$ is showed below

```{r example}
library(EstimationTools)

x <- rnorm(n = 10000, mean = 160, sd = 6)
fit <- maxlogL(x = x, link=list(over=c("sd"), fun=c("log.link")))
summary(fit)
```
